---
title: "My Prediction Assignment Writeup"
author: "Patricio Cofre"
date: "October 22, 2016"
output: html_document
---
### Summary
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

##### Libraries used
```{r, message=FALSE, warning=FALSE}
library(caret)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(rattle)
library(randomForest)
library(e1071)
```

###Loading data sets
```{r, message=FALSE, warning=FALSE}
training <- read.csv("pml-training.csv",na.strings=c("NA","#DIV/0!", ""))
testing <- read.csv("pml-testing.csv",na.strings=c("NA","#DIV/0!", ""))
```

*The training set consists of 19622 observations of 160 variables
*The testing set consists of 20 observations of 160 variables

### Cleaning data
Columns in the orignal training and testing datasets that are mostly filled with missing values are then removed count the number of missing values in each column of the full training dataset
```{r, message=FALSE, warning=FALSE}
training <-training[,colSums(is.na(training)) == 0]
testing <-testing[,colSums(is.na(testing)) == 0]
```

```{r, message=FALSE, warning=FALSE}
training   <-training[,-c(1:7)]
testing <-testing[,-c(1:7)]

dim ( training ) 
dim ( testing  )
```

Diagnoses predictors that have one unique value or predictors that are have both of the following characteristics
```{r}
ColumnsZVar <- nearZeroVar(training, saveMetrics = TRUE)
training <- training[, ColumnsZVar$nzv==FALSE]
training$classe = factor(training$classe)
```

Partitioning the training data This validation dataset will allow us to perform cross validation when developing our model.

### Partitioning the training data set to allow cross-validation
```{r}
set.seed(1234)
subTrain <- createDataPartition(y=training$classe, p=.75, list=FALSE)

TheTraining <- training[subTrain, ]
TheTesting <- training[-subTrain, ]
```

Dataset contains 59 variables, and the last column containing the 'class' variable we are trying to predict.

## Modelprediction 1 : Using Decision Tree
```{r, message=FALSE, warning=FALSE}
modelDT <- rpart(classe ~ ., data=TheTraining, method="class")
```

```{r, message=FALSE, warning=FALSE}
predictionDT <- predict(modelDT, TheTesting, type ="class")
```

### Plot of the Decision Tree
```{r, message=FALSE, warning=FALSE}
rpart.plot(modelDT, main="Decision Tree ", extra=102, under=TRUE, faclen=0)
fancyRpartPlot (modelDT, main="Decision Tree")
```

###Test results on our subTesting dataset
```{r}
confusionMatrix(predictionDT,TheTesting$classe)
```
The Confusion Matrix achieved  0.7394 % accuracy. 

## Modelprediction 2 : Using Random Forest

```{r}
modelRF <- randomForest(classe ~. , data=TheTraining, method="class")
print (modelRF)
```
### Predicting:

```{r}
predictionRF <- predict(modelRF, TheTesting, type = "class")
```

### Test results on subTesting dataset
```{r, echo=TRUE}
confusionMatrix(predictionRF, TheTesting$classe)
```

The Confusion Matrix achieved 99.51% accuracy.

###Decision & Conclusion

Random Forest algorithm performed better than Decision Trees.
Accuracy for Random Forest model was Accuracy 0.995, compared to 0.739 for Decision Tree model. 
Our Test data set comprises 20 cases. With an accuracy above 99% on our cross-validation data, we can expect that very few, or none, of the test samples will be 